---
title: "Customer Segmentation"
author: "Subash chander swaminathan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Import Dataset
```{r}
customer_df = read.csv("Mall_Customers.csv")
str(customer_df)
```


```{r}
names(customer_df)
```
```{r}
head(customer_df)
```
Customers Age
```{r}
summary(customer_df$Age)
```
```{r}
sd(customer_df$Age)
```
```{r}
hist(customer_df$Age,
     col="blue",
     main="Histogram to Show Count of Age Class",
     xlab="Age Class",
     ylab="Frequency",
     labels=TRUE)
boxplot(customer_df$Age,
        col="blue",
        main="Boxplot for Descriptive Analysis of Age")
```


Customers Annual Income
```{r}
summary(customer_df$Annual.Income..k..)
```
```{r}
sd(customer_df$Annual.Income..k..)
```
```{r}
hist(customer_df$Annual.Income..k..,
     col="#660033",
     main="Histogram for Annual Income",
     xlab="Annual Income Class",
     ylab="Frequency",
     labels=TRUE)
plot(density(customer_df$Annual.Income..k..),
     col="#660033",
     main="Density Plot for Annual Income",
     xlab="Annual Income Class",
     ylab="Density")
polygon(density(customer_df$Annual.Income..k..),
        col="#660033")
```


Customers spending score
```{r}
summary(customer_df$Spending.Score..1.100.)
```
```{r}
sd(customer_df$Spending.Score..1.100.)
```
```{r}
boxplot(customer_df$Spending.Score..1.100.,
        horizontal=TRUE,
        col="#6600cc",
        main="BoxPlot for Descriptive Analysis of Spending Score")
hist(customer_df$Spending.Score..1.100.,
     main="HistoGram for Spending Score",
     xlab="Spending Score Class",
     ylab="Frequency",
     col="#6600cc",
     labels=TRUE)
```


Number of Customer based on Gender
```{r}
a=table(customer_df$Gender)
barplot(a,main="Using BarPlot to display Gender Comparision",
        ylab="Count",
        xlab="Gender",
        col=rainbow(2),
        legend=rownames(a)
)


pct=round(a/sum(a)*100)
lbs=paste(c("Female","Male")," ",pct,"%",sep=" ")
library(plotrix)
pie3D(a,labels=lbs,
      main="Pie Chart Depicting Ratio of Female and Male")
```
Cluster model to segment customers
```{r}
library(purrr)
set.seed(1234)
```

Function to calculate total intra-cluster sum of square
```{r}
iss <- function(k) {
  kmeans(customer_df[,3:5],k,iter.max=100,nstart=100,algorithm="Lloyd" )$tot.withinss
}
k.values <- 1:10

iss_values <- map_dbl(k.values, iss)

plot(k.values, iss_values,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total intra-clusters sum of squares")
```
6 cluster combination produced better result that others
```{r}
library(cluster) 
library(gridExtra)
library(grid)


k2<-kmeans(customer_df[,3:5],2,iter.max=100,nstart=50,algorithm="Lloyd")
s2<-plot(silhouette(k2$cluster,dist(customer_df[,3:5],"euclidean")))

k3<-kmeans(customer_df[,3:5],3,iter.max=100,nstart=50,algorithm="Lloyd")
s3<-plot(silhouette(k3$cluster,dist(customer_df[,3:5],"euclidean")))

k4<-kmeans(customer_df[,3:5],4,iter.max=100,nstart=50,algorithm="Lloyd")
s4<-plot(silhouette(k4$cluster,dist(customer_df[,3:5],"euclidean")))

k5<-kmeans(customer_df[,3:5],5,iter.max=100,nstart=50,algorithm="Lloyd")
s5<-plot(silhouette(k5$cluster,dist(customer_df[,3:5],"euclidean")))

k6<-kmeans(customer_df[,3:5],6,iter.max=100,nstart=50,algorithm="Lloyd")
s6<-plot(silhouette(k6$cluster,dist(customer_df[,3:5],"euclidean")))

k7<-kmeans(customer_df[,3:5],7,iter.max=100,nstart=50,algorithm="Lloyd")
s7<-plot(silhouette(k7$cluster,dist(customer_df[,3:5],"euclidean")))

k8<-kmeans(customer_df[,3:5],8,iter.max=100,nstart=50,algorithm="Lloyd")
s8<-plot(silhouette(k8$cluster,dist(customer_df[,3:5],"euclidean")))

k9<-kmeans(customer_df[,3:5],9,iter.max=100,nstart=50,algorithm="Lloyd")
s9<-plot(silhouette(k9$cluster,dist(customer_df[,3:5],"euclidean")))

k10<-kmeans(customer_df[,3:5],10,iter.max=100,nstart=50,algorithm="Lloyd")
s10<-plot(silhouette(k10$cluster,dist(customer_df[,3:5],"euclidean")))
```
used fviz_nbclust() function to determine and visualize the optimal number of clusters

```{r}
library(NbClust)
library(factoextra)

fviz_nbclust(customer_df[,3:5], kmeans, method = "silhouette")
```
Used the clusGap() function for providing gap statistic as well as standard error
```{r}
set.seed(1234)
stat_gap <- clusGap(customer_df[,3:5], FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(stat_gap)
```

```{r}
k6<-kmeans(customer_df[,3:5],6,iter.max=100,nstart=50,algorithm="Lloyd")
k6
```

 Principal component analysis
```{r}
pcclust=prcomp(customer_df[,3:5],scale=FALSE)
summary(pcclust)
```


```{r}
pcclust$rotation[,1:2]
```

Visuvalized the clusters
```{r}
set.seed(1234)
ggplot(customer_df, aes(x =Annual.Income..k.., y = Spending.Score..1.100.)) + 
  geom_point(stat = "identity", aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
                       breaks=c("1", "2", "3", "4", "5","6"),
                       labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")

ggplot(customer_df, aes(x =Spending.Score..1.100., y =Age)) + 
  geom_point(stat = "identity", aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
                       breaks=c("1", "2", "3", "4", "5","6"),
                       labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")
kCols=function(vec){cols=rainbow (length (unique (vec)))
return (cols[as.numeric(as.factor(vec))])}

digCluster<-k6$cluster; dignm<-as.character(digCluster);  

plot(pcclust$x[,1:2], col =kCols(digCluster),pch =19,xlab ="K-means",ylab="classes")
legend("bottomleft",unique(dignm),fill=unique(kCols(digCluster)))
```
Cluster 1 and 4 – These two clusters consist of customers with medium PCA1 and medium PCA2 score.

Cluster 6 – This cluster represents customers having a high PCA2 and a low PCA1.

Cluster 5 – In this cluster, there are customers with a medium PCA1 and a low PCA2 score.

Cluster 3 – This cluster comprises of customers with a high PCA1 income and a high PCA2.

Cluster 2 – This comprises of customers with a high PCA2 and a medium annual spend of income.

